Equivalence Hypergraphs: E-Graphs for Monoidal Theories

### 2025-01-23 LICS_2025/monoidal_e_graphs/sections/conclusion.tex
wip

In the future,  we intend to investigate a number of natural extensions to mathematical treatment of (acyclic) e-graphs developed here.
Some of the extensions that can be considered are: support for `functorial boxes'~\cite{mellies_functorial_2006} (in addition to e-boxes), which have a variety of applications, including the Cartesian-closure required for the implementation of functional programming languages~\cite{ghica-zanassi2023string}
% , as sketched in Fig. \ref{fig:app1}(a)
; the `spiders' used by the ZX-calculus to model quantum circuits~\cite{coecke_interacting_2011,ZX}
% , as sketched in Fig. \ref{fig:app1}(b)
; or trace~\cite{joyal_geometry_1991, Hasegawa-traced} which can be used to express feedback in categorical models of digital circuits~\cite{ghica_jung_2017,ghica_compositional_2023} and, indeed in conventional e-graphs to encode infinite equivalence classes
% , as sketched in Fig.~\ref{fig:app2}. 
All these application domains stand to benefit from the use of optimisation techniques, which in turn can take advantage of the appropriate e-graph technique.

@@

In each case above,  we aim to combine our combinatorial representation with existing combinatorial representations of the relevant structure.  Indeed,  the hypergraph representation and DPO-rewriting theory of each case is already well-studied independently of any e-graph structure \cite{ghica_rewriting_2023,alvarez-picallo-functorial_2021}. % traced comonoid
%The hierarchical ``e-box'' structure within string diagrams and their associated e-hypergraphs (categorically modelled by semilattice enrichment) is essentially orthogonal to these developments. Combining the two, to give notions of e-hypergraphs specifically suited to these particular monoidal theories should be conceptually straightforward.  However, as we have seen in this paper, the technical details can be delicate. 
Finally,  a term calculus (e-terms) can be reverse engineered from the string-diagram notation~\cite{ghica_structural_nominal, Heijltjes:FMCII}. Although term calculi are less important in the algorithmics of optimisation, they still play a role in specification and in aiding readability.

>>>>

In the future,  we intend to investigate a number of natural extensions to mathematical treatment of (acyclic) e-graphs developed here.
Some of the extensions that can be considered are: support for `functorial boxes'~\cite{mellies_functorial_2006} (in addition to e-boxes), which have a variety of applications, including the Cartesian-closure required for the implementation of functional programming languages~\cite{ghica-zanassi2023string};
% , as sketched in Fig. \ref{fig:app1}(a)
the `spiders' used by the ZX-calculus to model quantum circuits~\cite{coecke_interacting_2011,ZX};
% , as sketched in Fig. \ref{fig:app1}(b)
or trace~\cite{joyal_geometry_1991, Hasegawa-traced} which can be used to express feedback in categorical models of digital circuits~\cite{ghica_jung_2017,ghica_compositional_2023} and, indeed in conventional e-graphs to encode infinite equivalence classes.
% , as sketched in Fig.~\ref{fig:app2}. 
All these application domains stand to benefit from the use of optimisation techniques, which in turn can take advantage of the appropriate e-graph technique.
In each case above,  we aim to combine our combinatorial representation with existing combinatorial representations of the relevant structure as the hypergraph representation and DPO-rewriting theory of each case is already well-studied independently of any e-graph structure \cite{ghica_rewriting_2023,alvarez-picallo-functorial_2021}.
 % traced comonoid

@@

% In each case above,  we aim to combine our combinatorial representation with existing combinatorial representations of the relevant structure.  Indeed,  the hypergraph representation and DPO-rewriting theory of each case is already well-studied independently of any e-graph structure \cite{ghica_rewriting_2023,alvarez-picallo-functorial_2021}. % traced comonoid
%The hierarchical ``e-box'' structure within string diagrams and their associated e-hypergraphs (categorically modelled by semilattice enrichment) is essentially orthogonal to these developments. Combining the two, to give notions of e-hypergraphs specifically suited to these particular monoidal theories should be conceptually straightforward.  However, as we have seen in this paper, the technical details can be delicate. 
% Finally,  a term calculus (e-terms) can be reverse engineered from the string-diagram notation~\cite{ghica_structural_nominal, Heijltjes:FMCII}. Although term calculi are less important in the algorithmics of optimisation, they still play a role in specification and in aiding readability.
